---
title: "Econometrics Session2 -- R Regressions"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---

Now we are trying to get ready for running regressions in R and for collecting nice output from the regressions. 


First, let's load (and potentially install) some packages.

```{r}
#install.packages("stargazer")
#install.packages("dplyr")
#install.packages("GGally")
#install.packages("ggplot2")
library(stargazer)
library(dplyr)
library(GGally)
library(ggplot2)

```




Next, we load data from the American Time Use Survey.

```{r}
# we want to read in the dataset "atus.csv", where all string variables are automatically transformed to factor variables
atus <- read.csv("atus.csv", stringsAsFactors = TRUE)
```

Here is a description of the variables in the data set: 

- `tuyear` survey year
- `occup_code` occupation code where individual works
- `ind_code` industry code where individual works
- `labor_status` labor force status 
- `ptft` whether the respondent works part-time or full-time 
- `work_class` class of work 
- `hourly_wage` hourly earnings at main job 
- `work_hrs_week` number of hours worked per week 
- `mult_jobs` had more than 1 jobs in 7 days prior to interview 
- `partner_hh` presence of respondent's spouse/unmarried partner in the household 
- `hh_size` number of people living in household 
- `hh_child` presence of household children aged < 18 



We would like to describe the variables that determine the hourly wage of an individual. To do so, we follow the 6 steps from class.

1. Look at the data

```{r}
#View(atus)
```

Anything that looks weird?

2. Summary statistics

```{r}
summary(atus)
```

Do you observe something strange?


Answer: Several Variables are not numeric. 
```{r}
str(atus)
```


```{r}
levels(atus$occup_code)
levels(atus$ind_code)
levels(atus$partner_hh)

```




3. Scatterplots and 4. Correlations


```{r}
ggpairs(atus[,c("hourly_wage", "work_hrs_week", "hh_size")])
```

5. Model

Let's discuss which variables should enter the model. [Open discussion]


6. Estimate and Assess Model

[Use variables that you determined in the discussion.]
```{r}
model1 <- lm(hourly_wage ~ hh_size + work_hrs_week + occup_code, data = atus)

summary(model1)
```

Does the model make sense? Are the signs and significance levels as you expected? Interpret the coefficients.

Look at the adjusted R2. 


Now let's learn about a packages that simplifies the comparison of different models and helps to produce nice regression output.

```{r}

# for summary statistics 
stargazer(atus, type = 'text', digits = 1)

#for regression output
stargazer(model1, type = 'text')

```



You can experiment with different model specifications. Does the R2 change? Do the coefficients change? What is your preferred model specification? 


```{r}
model2 <- lm(hourly_wage ~ hh_size + work_hrs_week + ptft , data = atus)
model3 <- lm(hourly_wage ~ hh_size + work_hrs_week + ptft + occup_code, data = atus)
stargazer(model1, model2, model3, type = 'text')

```




# Regression formulas:

Let's learn about additional options how to specify regression models. 

- `outcome ~ independent_variable1 + independent_variable2`
- factor variables are automatically turned into dummies `y ~ x + factor(z)`  (We have seen this with the `occup_code`)
- interactions: `y ~ x*z` 
- only interactions (not individual): `y ~ x:z`
- full interaction sets: `y ~ (x1 + x2 + x3)^2`
- can use functions: `ln(y) ~ x` 
- Use `I()` if you want to do calculations with variables first: `I(y == 1) ~ I(x^2)`
- With many `x` variables can abbreviate `y ~ .`


Exercise: Try several alternative specifications: log hourly wage, interactions between `ptft` and working hours, etc...
How are the coefficients interpreted?

# Prediction 

Finally, we can make a prediction for new data using our preferred model. Let us assume, we have an individual with 

```{r}

model_final <- lm(hourly_wage ~ hh_size + work_hrs_week + ptft + occup_code , data = atus)


new_data <- data.frame(hh_size = 3, work_hrs_week = 40, occup_code = "professional", ptft = "FT")

predicted_hourly_wage <- predict(model_final, newdata = new_data)

```


