---
title: "Econometrics Session1 -- R Introduction"
output:
  html_document: default
  html_notebook: default
---

# Introduction to R

## 0: Some Basics 

- windows in Rstudio:
    - The upper-left window is the script. a text file where you can type commands and run the commands, add comments and save your code for later use. 
    - The lower-left window is the console. Here is where our results will show and (hopefully few) error messages. you can also directly type commands in this window.
    - The upper-right window provides an overview the data you have in your working environment.
    - The lower-right window is multifunctional. There are multiple tabs, such as files, plots, packages, help and viewer. Especially the tabs plots, packages and help are very useful as we will see soon.


### getting help
- type `?command` or `help(command)` to find out more! the R documentation will show up.
```{r}
help(setwd)

```


### executing the code from the script
- Windows: `ctrl`+`enter`
- Mac: `command`+`enter`


## 1: Set your Working Directory

The first step is setting the working directory which is basically specifying the path of the location from which you want to read files and to which you want to write files in R.   

Insert your folder path between the `""`. 
```{r}
#setwd("/home/jan/Dropbox/Teaching/Econometrics/Econometrics MIM 2021/Session1_Overview")
getwd()
```
A better alternative is to create a new project which lives in its own folder. This allows you to share code with others without changing the file path. You can *relative* rather than *absolute* file paths.

To check what is the current working directory:


```{r}
getwd()
```

##  2: Data Types 

Use the `class()` function to find out what type an object is.

### 1. Character/string: 

Each element in the object is a string of one or more characters.

Lets see an example:
We will create a vector with string elements. A vector is an array of data elements of the same data type.
Use the `c()` function to combine elements in a vector.
To assign the vector to a variable name we use the assign operator `<-`: `name <- c()`

Strings have to be written within `""`

Exercise: assign vector to name fruits

```{r}
fruits<-c("banana", "apple", "pear", "mango")
class(fruits)
is.vector(fruits)

```


### 2. Factor 
Factor variables contain a set of numeric codes with character-valued levels.  
Let's create one for the gender of a group of 8 studens, of whom 3 are female and 5 are male.

```{r}
gender_students<-factor(c("female", "male", "female", "male", "female", "male", "male", "male"))
levels(gender_students)
nlevels(gender_students)

```


### 3. Numeric

```{r}
class(2.3)
is.numeric(2.3)

```




### converting data types

```{r}
as.character(6)
as.numeric("6.5")
as.integer("6.5")
as.integer(6.7)
as.numeric(fruits)
fruits2<-as.factor(fruits)
levels(fruits2)
as.numeric(fruits2)

```


## 3: Loading Data into R

Clean work space:

```{r}
rm(list =ls())

```


We are going to load the california test score data into R. This is a csv file. The data will be loaded in a data frame.  
A data frame is a tightly coupled collection of variables with the same number of rows. A data frame can contain columns with different classes.  

Read csv file and save in dataframe called `mydata`

```{r}
mydata<-read.csv(file="caschool_adapted.csv", header=TRUE, sep=",")

```

What is in the dataset?
```{r}
View(mydata)
```


- `DIST_CODE`:		DISTRICT CODE;
- `READ_SCR`:		AVG READING SCORE;
- `MATH_SCR`:		AVG MATH SCORE;
- `COUNTY`:		COUNTY;
- `DISTRICT`:		DISTRICT;
- `GR_SPAN`:		GRADE SPAN OF DISTRICT;
- `ENRL_TOT`:		TOTAL ENROLLMENT;
- `TEACHERS`:		NUMBER OF TEACHERS;
- `COMPUTER`:		NUMBER OF COMPUTERS;
- `TESTSCR`:		AVG TEST SCORE (= (READ_SCR+MATH_SCR)/2 );
- `EXPN_STU`:		EXPENTITURES PER STUDENT ($’S);
- `EL_PCT`:		PERCENT OF ENGLISH LEARNERS;
- `MEAL_PCT`:		PERCENT QUALIFYING FOR REDUCED-PRICE LUNCH;
- `CALW_PCT`:		PERCENT QUALIFYING FOR CALWORKS; (public assistance program that provides cash aid to families)
- `AVGINC`:		DISTRICT AVERAGE INCOME (IN $1000'S);
                                  

## 4: Exploring the Data  

```{r}
summary(mydata) # Provides basic descriptive statistics and frequencies.
sapply(mydata, class) # provides class of each variable
sapply(mydata, mean) # provides the mean
str(mydata) # Provides the structure of the dataset
names(mydata) # Lists variables in the dataset

```



### Selecting a subset of the data

If you want to select a subset of the data from the dataframe, you use `[,]` brackets.
Fill in the row(s) and column(s) you want: `[rows, columns]` using the colon, e.g., `1:10` gives you elements 1 to 10.

```{r}
mydata[1:10,1:4]
mydata[2,6] # cell
mydata[6, ] # all columns
```

### Accessing variables in data frame:
`mean(computer)` does not work! R cannot find the object.
```{r}
mean(mydata$computer) # need to specify dataframe named "data" in which the variable computer is stored
summary(mydata$avginc)

```


One factor variable is `gr_span`. It indicates the grade span of the district. We can create a frequency table with the `table()` command.  


```{r}
t = table(mydata$gr_span)  #first we create the table and save it in t
print(t) #then we print t to see the results

```

Do we have missing observations?

```{r}
colSums(is.na(mydata)) # no we don't

```


We can look at missing observations of a specific variable
```{r}
sum(is.na(mydata$computer))

```

### Installing and loading packages


R provides a ton of packages that add specific functionality to the base R. 
```{r}
#install.packages("psych")
library(psych)

```

Functions provided: 
- `trimmed`: trimmed mean (trim defaulting to 0.1)
- `mad`: median of the absolute deviation (from the median) 
```{r}
describe(mydata)

```

### Correlations
We can construct a table with correlations for the numeric variables.

```{r}
correlations <- cor(mydata[,8:16]) 
print(correlations)
correlations2<-cor(mydata[, c("expn_stu", "avginc", "math_scr")])
print(correlations2)
```

Correlations between two variables:
```{r}
cor(mydata$expn_stu, mydata$avginc)

```



## 5: Constructing some Variables 
Let's construct the student teacher ratio, `STR`:	STUDENT TEACHER RATIO (ENRL_TOT/TEACHERS);

```{r}
mydata$str<-(mydata$enrl_tot/mydata$teachers)
summary(mydata$str)

```




### Exercise: 
Construct the computer student ratio.

```{r}
mydata$csr<-(mydata$computer/mydata$enrl_tot)
summary(mydata$csr)
```



## 6: Creating Indicator Variables
It might be useful to create an indicator variable.
Here is an example: 
- we will create a dummy variable that equals 1 if testscore is above 650 and 0 otherwise
- we will save this variable in our data frame `mydata` and name it `above_dummy`

```{r}
mydata$above_dummy<-as.numeric(mydata$testscr>650) 

```

We might want to add a variable label to this variable that describes the variable. For this we need the Hmisc package.

```{r}
install.packages("Hmisc")
library(Hmisc)
label(mydata$above_dummy) <- " Indicator variable for test scores above 650" 
summary(mydata$above_dummy) #Gmd is gini difference 

```

We can again create a frequency table

```{r}
t = table(mydata$above_dummy) 
print(t)

```



## 7: Plots

Simple histogram:
```{r}
hist(mydata$math_scr)
?hist

```

More options in histogram:
```{r}

hist(mydata$math_scr, 
     main="Histogram of Math Scores",  	# definition of the title
     col = "light blue", 			# changing color of the bins
     breaks=seq(600,710,5), # changing mmidpoints of the bins
      xlab="Match Scores", 
     ylab= "Frequency")			

```


Simple scatterplot:
```{r}
plot(mydata$str, mydata$math_scr, main="Scatterplot",
     xlab="student-teacher ratio",
     ylab="Math scores")

plot(mydata$avginc, mydata$math_scr, main="Scatterplot",
     xlab="Average income",
     ylab="Math scores")

```



### Exercise: 
Produce some plots for computer per student ratio and Math scores and for average income and read scores.




```{r}
```







## 8: Saving Data

Write the data to your working directory, save in a csv file.
```{r}
write.table(mydata, file = "caschool.txt", sep = "\t")

```



## 9: Extra Exercises
Depending on time, we can do some more exercises.

1. What is the correlation between avg income in a district and the student teacher ratio?

```{r}

```



2. Create an indicator, that equals 1 for observations with expenditures per student (expn_stu) above 5000. Name the indicator `expn_dummy`.

```{r}



```






3. Label the variable `expn_dummy`, describe that it is an indicator variable for expenditures per student above 5000.

```{r}

```





4. plot a histogram of average income. make the columns have the color "red" and set the bins to a sequence from 0 to 60, with bin width of 5 `(breaks=seq(0,60,5))`.


```{r}


```



