---
title: "Econometrics Session4 -- R Regressions"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---

1. Prep work
```{r}
library(stargazer)
library(dplyr)
library(GGally)
library(ggplot2)
```

```{r}
hollywood <- read.csv("hollywood.csv")
```

```{r}
#View(hollywood)
```

2.Constructing some Variables 
Converting seuquel to dummy variable, `sequeldummy`
Construct the return ratio, `roi`: (openinggross/budget)
interaction term opensummer x openholiday, `sumhol`: (opensummer*openholiday)
interaction term criticsopinion x sequel, `criseq`: (criticsopinion*sequeldummy)
polynomial of degree 2 for openingtheaters, `openingtheaters2`: (openingtheaters^2)

```{r}
hollywood$sequeldummy <- as.numeric(hollywood$sequel=="yes")
hollywood$roi <- (hollywood$openinggross/hollywood$budget)
hollywood$sumhol <- (hollywood$opensummer*hollywood$openholiday)
hollywood$criseq <- (hollywood$criticsopinion*hollywood$sequeldummy)
hollywood$openingtheaters2 <- (hollywood$openingtheaters)^2

#summary(hollywood$roi)
```

3. Scatterplots and Correlations
```{r}
#ggp <- ggpairs(hollywood[,c("openinggross", "budget", "openingtheaters", "criticsopinion", "sumhol", "criseq")])
ggp <- ggpairs(hollywood[,c("opensummer", "openholiday", "sumhol", "criticsopinion", "sequeldummy", "criseq")])
print(ggp, progress = FALSE)
```

4. Estimate and Assess Model
```{r}
model1 <- lm(openinggross ~ budget + openingtheaters + knownstory + sequeldummy + opensummer + openholiday + openchristmas + criticsopinion + comedy, data = hollywood)
model2 <- lm(openinggross ~ budget + openingtheaters + knownstory + opensummer + openholiday + openchristmas + criticsopinion + comedy, data = hollywood) #w/o sequel
model3 <- lm(openinggross ~ budget + openingtheaters + knownstory + sequeldummy + opensummer + openholiday + openchristmas + comedy, data = hollywood) #w/o criticsopinion
model4 <- lm(openinggross ~ budget + openingtheaters + opensummer + openholiday + openchristmas + comedy, data = hollywood) #w/o sequel or criticsopinion or knownstory
model4 <- lm(openinggross ~ budget + sequeldummy + criticsopinion + openingtheaters, data = hollywood) #w/o  knownstory + opensummer + openholiday + openchristmas + comedy
model_wo_dummy <- lm(openinggross ~ budget + sequel + criticsopinion + openingtheaters, data = hollywood) #w/o  knownstory + opensummer + openholiday + openchristmas + comedy

stargazer(model1, model2, model3, model4, model_wo_dummy, type = 'text')
```


4. Estimate and Assess New Models
```{r}
model4 <- lm(openinggross ~ budget + sequel + criticsopinion + openingtheaters, data = hollywood) #w/o  knownstory + opensummer + openholiday + openchristmas + comedy
model5 <- lm(openinggross ~ budget + criticsopinion + openingtheaters + sumhol + criseq + openingtheaters2, data = hollywood) #w/o  knownstory + opensummer + openholiday + openchristmas + comedy


stargazer(model4, model5, type = 'text')
```

# Prediction 

Finally, we can make a prediction for new data using our preferred model. Let us assume, we have an individual with 

```{r}
#no critics opinion
model_newmovie <- lm(openinggross ~ budget + openingtheaters + knownstory + sequeldummy + opensummer + openholiday + openchristmas + comedy, data = hollywood)

newmovie_data <- data.frame(budget = 40000000, openingtheaters = 2800, knownstory = 1, sequeldummy = 0, opensummer = 1, openholiday = 0, openchristmas = 0, comedy = 1)

predicted_gross <- predict(model_newmovie, newdata = newmovie_data)

predicted_gross/1000000

#how to do confidence interval??
```


