---
title: "R Notebook"
output: html_notebook
---

# Load data

```{r, results=FALSE, message=FALSE}


library(AER)
library(stargazer)
library(ivpack)
library(dplyr)

```


# Generate new variables

```{r}

cigs = read.csv("cigs_data.csv")

cigs <- mutate(cigs,
               #deflate by CPI to get real prices
               rprice = price / cpi,
               rincome = (income/population)/cpi,
               #log variables
               lprice = log(rprice),
               lquant = log(packs),
               lincome = log(rincome))
```



# Summary statistics
```{r}
stargazer(cigs, type="text", median=TRUE,
          digits=2, title="Cigarette data 1995")
```

# Estimate OLS

```{r, warning=FALSE}
ols <- lm(lquant ~ lprice, data = cigs)
stargazer(ols,
          title="OLS Regression", type = 'text', df = FALSE, digits = 5)
```



# IV regression - First Stage

Regress $X$ on $Z$ (and any other exogenous variables $W$ if we had them)

```{r, warning=FALSE}
cigs <- mutate(cigs, tdiff = (taxs-tax)/cpi)

first1 <- lm(lprice ~ tdiff, data = cigs) 
stargazer(first1,  
          title="First-stage Regression", type="text", 
          df=FALSE, digits=5)
```

# IV regression - Second Stage

Manually run the second stage using predicted values from first stage.
```{r,warning=FALSE}
cigs$lpricehat <- fitted(first1)

second1 <- lm(lquant ~ lpricehat, data = cigs)
stargazer(second1,  
          title="Second Stage Regression", type="text", 
          df=FALSE, digits=5)
```


```{r, warning=FALSE}
iv1 = ivreg(lquant ~ lprice | tdiff , data = cigs)
stargazer(iv1, 
          title="IV Regression", type="text", 
          df=FALSE, digits=5)
```


# Estimated demand equation

$$\ln(Q_i^{cigarettes})=9.72-1.08\ln(P_i^{cigarettes})$$


# General IV Model

$$Y_i=\beta_0+\beta_1X_{1i}+...+\beta_kX_{ki}+\beta_{k+1}W_{1i}+...+\beta_{k+r}W_{ri}+u_i$$



# General IV Model in our example

We can add a control variable (`lincome`). 

Suppose now, that we have two instruments:
- Cigarette tax `tdiff`
- General tax in real terms (`tax/cpi`).


# General IV Model in our example

```{r, warning = FALSE}
#Estimate the first stage with all exogenous regressors, i.e. control vars and instruments
first2 = lm(lprice ~ lincome + tdiff + I(tax/cpi), data = cigs)
stargazer(first2, 
          title="First Stage with 3 Exogenous Regressors", 
          type="text", df=FALSE, digits=5) 
```

# Use `ivreg` command

```{r, warning=FALSE}
iv2 = ivreg(lquant ~ lprice + lincome | lincome + tdiff + I(tax/cpi),
            data = cigs)
stargazer(iv2, 
          title="IV Regression", type="text", 
          df=FALSE, digits=5) 

```


```{r}
first1 = lm(lprice ~ tdiff, data = cigs)
summary(first1)
```


# Your turn: Check instrument relevance for the second instrument `tax/cpi`
```{r}


```

# Better: Do this in the general model

- $H_0: \beta_{tdiff}=\beta_{tax/cpi}=0$ in your first stage.
- Decision rule: reject $H_0$ if $F>F^*$
- Rule of thumb: conclude that you have relevant instruments if $F>10$
- In practice: conclude that you have relevant instruments if $p<0.05$

# Results:
```{r}
summary(iv2, diagnostics = TRUE)
```














Decision: reject $H_0$, conclude that you have relevant instruments.



# Checking instrument exogeneity: Your task

Manually compare how different the estimates are if you use the two instruments in separate iv regressions.
\tiny
```{r, warning=FALSE}

```

# Test of overidentifying restrictions (instrument exogeneity)

A \textcolor{red}{test of overidentifying} restrictions makes this comparison in a statistically precise way.

- Warning: You need at least two exogenous instruments to check for this here!
- $H_0:$ All instruments are exogenous
- Decision rule: reject $H_0$ if Sargan statistic $>$ critical value
- In practice: reject $H_0$ that instruments are exogenous if $p<0.05$ (Note: this is bad news! You \textit{want} your instruments to be exogenous!)


# Results

```{r}
```


# Overview on all results

```{r, warning=FALSE}
stargazer(ols, first1, iv1, first2, iv2, 
          title="All Regression Output", type="text", 
          df=FALSE, digits=5) 

```


# Second Overview
```{r, warning=FALSE}
stargazer(ols, first1, iv1, iv2, 
          title="IV Regression", type="text", 
          df=FALSE, digits=5) 

```


